---
title: "Gambia Min_Max Temperatures"
author: "Orero Levi"
date: "10/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load required packages }

library(raster)
library(ggplot2)
library(leaflet)
library(dplyr)
library(lubridate)
library(tidyverse)
library(sf)
```

```{r read the data}
#raster <- getData(name = 'worldclim', var = 'tmin', res =10, 
                  #lat= 13.451290, lon =-16.573000)## land surface temperature(lst) raster

raster2 <- brick('wc2.1_2.5m_prec_2016-09.tif')
raster3 <- brick('gam_climate/tmax/wc2.1_2.5m_tmax_2000-01.tif')

# Read shapefile method 1
ward <- shapefile("Gmb_adm0.shp")## 


ward2 = ward

ward3 = ward

wardcoun <- getData(name = 'GADM', country = "GMB", level = 1)


#ward <- spTransform(ward,crs(raster)) ## transform wards


# Raster 2
ward2 <- spTransform(ward2,crs(raster2)) ## transform wards

# Raster 3
ward3 <- spTransform(ward3,crs(raster3)) ## transform wards
```


```{r read data for regions}
# Method 1
gam_regions <- shapefile("gmb_adm1/gmb_admbnda_adm1_lga_1m_wfp_ocha.shp")

# Method 2
gam_regions2 <- st_read("GAM.shp")

```





```{r check the data}
raster
raster2
names(raster)
names(raster2)
names(raster) <- paste0(rep(month.abb, 2 ), sort(rep(2000:2018, 12)))

names(raster3)
```

```{r}
#plot(raster[[1]]) + plot(ward, add = T)

#image <- crop(tab, wardcoun)
#plot(image)
```

```{r crop the tif to just gambia}
#e_x_t <- as(extent(-16.81736, -13.79093, 13.06466, 13.82689), 'SpatialPolygons')
#crs(e_x_t) <- crs(raster)
#cropd <- crop(raster, e_x_t)

# Raster 2
e_x_t2 <- as(extent(-16.81736, -13.79093, 13.06466, 13.82689), 'SpatialPolygons')
crs(e_x_t2) <- crs(raster2)
cropd2 <- crop(raster2, e_x_t2)


# Raster 3
e_x_t3 <- as(extent(-16.81736, -13.79093, 13.06466, 13.82689), 'SpatialPolygons')
crs(e_x_t3) <- crs(raster3)
cropd3 <- crop(raster3, e_x_t3)
```

```{r plot the cropped raster}
#plot(cropd[[1]]) + plot(ward, add = T)

# Raster 2
plot(cropd2[[1]]) + plot(ward2, add = T)

# Raster 3
plot(cropd3[[1]]) + plot(ward3, add = T)

title(main = 'Maximum Temperature (\u00B0C) in the Gambia - January 2000')
```
```{r}
plot(mask(gam_2000_2018_tmax_st, gam_regions)[[13]]) + plot(ward, add = T)
```

## Function to plot and save monthly max temperatures
```{r}
mapper <- function(x,y,z){ #x is raster stack, #y is mask parameter # z is region
  for (i in 1:12){
    range = seq(i+12,204,12)
    #mean of layers by month
    layer = mean(x[[range]])
    # save image
    tiff( filename = paste0("gambia_plot ", month.name[i], ".tiff"), width= 850,
          height = 600,units = "px")
    #plot shapefile and raster stack
    plot(mask(layer, y)) + plot(z, add = T)
    title(main = paste0('Maximum Temperature (\u00B0C) in the Gambia in ', month.name[i]))
    dev.off()
  } 
}
```


```{r}
mapper(gam_2000_2018_tmax_st, gam_regions, ward)

# Patch into one figure 2001-2016
tiff( filename = "gambia_plot.tiff", width= 850,
          height = 850,units = "px")
par(mfrow = c(4,3), oma = c(0, 0, 2, 1))
plot(mask(mean(gam_2000_2018_tmax_st[[seq(13,204,12)]]), gam_regions)) + plot(ward, add = T)
title(main = month.name[1])
plot(mask(mean(gam_2000_2018_tmax_st[[seq(14,204,12)]]), gam_regions)) + plot(ward, add = T)
title(main = month.name[2])
plot(mask(mean(gam_2000_2018_tmax_st[[seq(15,204,12)]]), gam_regions)) + plot(ward, add = T)
title(main = month.name[3])
plot(mask(mean(gam_2000_2018_tmax_st[[seq(16,204,12)]]), gam_regions)) + plot(ward, add = T)
title(main = month.name[4])
plot(mask(mean(gam_2000_2018_tmax_st[[seq(17,204,12)]]), gam_regions)) + plot(ward, add = T)
title(main = month.name[5])
plot(mask(mean(gam_2000_2018_tmax_st[[seq(18,204,12)]]), gam_regions)) + plot(ward, add = T)
title(main = month.name[6])
plot(mask(mean(gam_2000_2018_tmax_st[[seq(19,204,12)]]), gam_regions)) + plot(ward, add = T)
title(main = month.name[7])
plot(mask(mean(gam_2000_2018_tmax_st[[seq(20,204,12)]]), gam_regions)) + plot(ward, add = T)
title(main = month.name[8])
plot(mask(mean(gam_2000_2018_tmax_st[[seq(21,204,12)]]), gam_regions)) + plot(ward, add = T)
title(main = month.name[9])
plot(mask(mean(gam_2000_2018_tmax_st[[seq(22,204,12)]]), gam_regions)) + plot(ward, add = T)
title(main = month.name[10])
plot(mask(mean(gam_2000_2018_tmax_st[[seq(23,204,12)]]), gam_regions)) + plot(ward, add = T)
title(main = month.name[11])
plot(mask(mean(gam_2000_2018_tmax_st[[seq(24,204,12)]]), gam_regions)) + plot(ward, add = T)
title(main = month.name[12])
mtext('Maximum Temperature (\u00B0C) in the Gambia 2001-2016', outer = TRUE, cex = 1.5)

tiff( filename = "gambia_plot.tiff", width= 850,
          height = 850,units = "px")
```

```{r plotTemperatures_max}
# Patch into one figure 2001-2018
tiff( filename = "gambia_plot_01-18.tiff", width= 850,
          height = 850,units = "px")
par(mfrow = c(4,3), oma = c(0, 0, 2, 1))
plot(mask(mean(gam_2000_2018_tmax_st[[seq(13,228,12)]]), gam_regions)) + plot(ward, add = T)
title(main = month.name[1])
plot(mask(mean(gam_2000_2018_tmax_st[[seq(14,228,12)]]), gam_regions)) + plot(ward, add = T)
title(main = month.name[2]) 
plot(mask(mean(gam_2000_2018_tmax_st[[seq(15,228,12)]]), gam_regions)) + plot(ward, add = T)
title(main = month.name[3])
plot(mask(mean(gam_2000_2018_tmax_st[[seq(16,228,12)]]), gam_regions)) + plot(ward, add = T)
title(main = month.name[4])
plot(mask(mean(gam_2000_2018_tmax_st[[seq(17,228,12)]]), gam_regions)) + plot(ward, add = T)
title(main = month.name[5])
plot(mask(mean(gam_2000_2018_tmax_st[[seq(18,228,12)]]), gam_regions)) + plot(ward, add = T)
title(main = month.name[6])
plot(mask(mean(gam_2000_2018_tmax_st[[seq(19,228,12)]]), gam_regions)) + plot(ward, add = T)
title(main = month.name[7])
plot(mask(mean(gam_2000_2018_tmax_st[[seq(20,228,12)]]), gam_regions)) + plot(ward, add = T)
title(main = month.name[8])
plot(mask(mean(gam_2000_2018_tmax_st[[seq(21,228,12)]]), gam_regions)) + plot(ward, add = T)
title(main = month.name[9])
plot(mask(mean(gam_2000_2018_tmax_st[[seq(22,228,12)]]), gam_regions)) + plot(ward, add = T)
title(main = month.name[10])
plot(mask(mean(gam_2000_2018_tmax_st[[seq(23,228,12)]]), gam_regions)) + plot(ward, add = T)
title(main = month.name[11])
plot(mask(mean(gam_2000_2018_tmax_st[[seq(24,228,12)]]), gam_regions)) + plot(ward, add = T)
title(main = month.name[12])
mtext('Maximum Temperature (\u00B0C) in the Gambia 2001-2018', outer = TRUE, cex = 1.5)
```



```{r}
# Plot country shapefile
leaflet()%>%
  addTiles()%>%
  addPolygons(data = ward)

# Plot regions shapefile
leaflet()%>%
  addTiles()%>%
  addPolygons(data = gam_regions)

# Plot filtered shapefiles
gam_regions_lower <- gam_regions2 %>% 
  filter(ADM1 == "Lower River")

gam_regions_upper <- gam_regions2 %>% 
  filter(ADM1 == "Upper River")

gam_regions_central <- gam_regions2 %>% 
  filter(ADM1 == "Central River")

leaflet()%>%
  addTiles()%>%
  addPolygons(data = gam_regions_lower)
```


```{r}
# Raster 3
plot(cropd3[[1]]) + plot(gam_regions_lower, add = T)
```

```{r masking only gambia}
mr <- mask(cropd3, gam_regions)
plot(mr)
```


```{r extract values}

# Raster 2
extract2 <- raster::extract(cropd2, ward2) %>%
  data.frame()

head(extract2)
dim(extract2)

# Raster 3
extract3 <- raster::extract(cropd3, ward3) %>%
  data.frame()

head(extract3)
dim(extract3)
```



























```{r read all files and stack them}
files <- list.files(path="C:/Users/LORERO/OneDrive - CGIAR/Documents/Gambia-Fire/gam_climate/tmax", pattern="*.tif", full.names=TRUE, recursive=FALSE)

# Create function
maskerTX <- function(x){
  t <- brick(x, header=TRUE) # load file
  # crop the tif to just gambia
  e_x_t <- as(extent(-16.81736, -13.79093, 13.06466, 13.82689), 'SpatialPolygons')
  crs(e_x_t) <- crs(t)
  cropd <- crop(t, e_x_t)
  # mask only gambia 
  mr <- mask(cropd, gam_regions)
  
  # write to file
  #save(mr, file = paste0(x,"saved.tif"))
}


#################
# Stack 2000 values

gam_2000_tmax <- lapply(files[1:12],maskerTX)
gam_2000_tmax_st <- stack(gam_2000_tmax)

# give proper names
names(gam_2000_tmax_st) <- paste("tmax", outer(month.abb, 2000, paste, sep="-"), sep="-")
names(gam_2000_tmax_st)
```
# Stack 2001 values
```{r}
gam_2001_tmax <- lapply(files[13:24],maskerTX)
gam_2001_tmax_st <- stack(gam_2001_tmax)

# give proper names
names(gam_2001_tmax_st) <- paste("tmax", outer(month.abb, 2001, paste, sep="-"), sep="-")
names(gam_2001_tmax_st)
```

# Stack 2002 values
```{r}
gam_2002_tmax <- lapply(files[25:36],maskerTX)
gam_2002_tmax_st <- stack(gam_2002_tmax)

# give proper names
names(gam_2002_tmax_st) <- paste("tmax", outer(month.abb, 2002, paste, sep="-"), sep="-")
names(gam_2002_tmax_st)
```

# Stack 2003 values
```{r}
gam_2003_tmax <- lapply(files[37:48],maskerTX)
gam_2003_tmax_st <- stack(gam_2003_tmax)

# give proper names
names(gam_2003_tmax_st) <- paste("tmax", outer(month.abb, 2003, paste, sep="-"), sep="-")
names(gam_2003_tmax_st)
```

# Stack 2004 values
```{r}
gam_2004_tmax <- lapply(files[49:60],maskerTX)
gam_2004_tmax_st <- stack(gam_2004_tmax)

# give proper names
names(gam_2004_tmax_st) <- paste("tmax", outer(month.abb, 2004, paste, sep="-"), sep="-")
names(gam_2004_tmax_st)
```


# Stack 2005 values
```{r}
gam_2005_tmax <- lapply(files[61:72],maskerTX)
gam_2005_tmax_st <- stack(gam_2005_tmax)

# give proper names
names(gam_2005_tmax_st) <- paste("tmax", outer(month.abb, 2005, paste, sep="-"), sep="-")
names(gam_2005_tmax_st)
```

# Stack 2006 values
```{r}
gam_2006_tmax <- lapply(files[73:84],maskerTX)
gam_2006_tmax_st <- stack(gam_2006_tmax)

# give proper names
names(gam_2006_tmax_st) <- paste("tmax", outer(month.abb, 2006, paste, sep="-"), sep="-")
names(gam_2006_tmax_st)
```

# Stack 2007 values
```{r}
gam_2007_tmax <- lapply(files[85:96],maskerTX)
gam_2007_tmax_st <- stack(gam_2007_tmax)

# give proper names
names(gam_2007_tmax_st) <- paste("tmax", outer(month.abb, 2007, paste, sep="-"), sep="-")
names(gam_2007_tmax_st)
```

# Stack 2008 values
```{r}
gam_2008_tmax <- lapply(files[97:108],maskerTX)
gam_2008_tmax_st <- stack(gam_2008_tmax)

# give proper names
names(gam_2008_tmax_st) <- paste("tmax", outer(month.abb, 2008, paste, sep="-"), sep="-")
names(gam_2008_tmax_st)
```


# Stack 2009 values
```{r}
gam_2009_tmax <- lapply(files[109:120],maskerTX)
gam_2009_tmax_st <- stack(gam_2009_tmax)

# give proper names
names(gam_2009_tmax_st) <- paste("tmax", outer(month.abb, 2009, paste, sep="-"), sep="-")
names(gam_2009_tmax_st)
```


# Stack 2010 values
```{r}
gam_2010_tmax <- lapply(files[121:132],maskerTX)
gam_2010_tmax_st <- stack(gam_2010_tmax)

# give proper names
names(gam_2010_tmax_st) <- paste("tmax", outer(month.abb, 2010, paste, sep="-"), sep="-")
names(gam_2010_tmax_st)
```


# Stack 2011 values
```{r}
gam_2011_tmax <- lapply(files[133:144],maskerTX)
gam_2011_tmax_st <- stack(gam_2011_tmax)

# give proper names
names(gam_2011_tmax_st) <- paste("tmax", outer(month.abb, 2011, paste, sep="-"), sep="-")
names(gam_2011_tmax_st)
```


# Stack 2012 values
```{r}
gam_2012_tmax <- lapply(files[145:156],maskerTX)
gam_2012_tmax_st <- stack(gam_2012_tmax)

# give proper names
names(gam_2012_tmax_st) <- paste("tmax", outer(month.abb, 2012, paste, sep="-"), sep="-")
names(gam_2012_tmax_st)
```


# Stack 2013 values
```{r}
gam_2013_tmax <- lapply(files[157:168],maskerTX)
gam_2013_tmax_st <- stack(gam_2013_tmax)

# give proper names
names(gam_2013_tmax_st) <- paste("tmax", outer(month.abb, 2013, paste, sep="-"), sep="-")
names(gam_2013_tmax_st)
```

# Stack 2014 values
```{r}
gam_2014_tmax <- lapply(files[169:180],maskerTX)
gam_2014_tmax_st <- stack(gam_2014_tmax)

# give proper names
names(gam_2014_tmax_st) <- paste("tmax", outer(month.abb, 2014, paste, sep="-"), sep="-")
names(gam_2014_tmax_st)
```


# Stack 2015 values
```{r}
gam_2015_tmax <- lapply(files[181:192],maskerTX)
gam_2015_tmax_st <- stack(gam_2015_tmax)

# give proper names
names(gam_2015_tmax_st) <- paste("tmax", outer(month.abb, 2015, paste, sep="-"), sep="-")
names(gam_2015_tmax_st)
```


# Stack 2016 values
```{r}
gam_2016_tmax <- lapply(files[193:204],maskerTX)
gam_2016_tmax_st <- stack(gam_2016_tmax)

# give proper names
names(gam_2016_tmax_st) <- paste("tmax", outer(month.abb, 2016, paste, sep="-"), sep="-")
names(gam_2016_tmax_st)
```

# Stack 2017 values
```{r}
gam_2017_tmax <- lapply(files[205:216],maskerTX)
gam_2017_tmax_st <- stack(gam_2017_tmax)

# give proper names
names(gam_2017_tmax_st) <- paste("tmax", outer(month.abb, 2017, paste, sep="-"), sep="-")
names(gam_2017_tmax_st)
```


# Stack 2018 values
```{r}
gam_2018_tmax <- lapply(files[217:228],maskerTX)
gam_2018_tmax_st <- stack(gam_2018_tmax)

# give proper names
names(gam_2018_tmax_st) <- paste("tmax", outer(month.abb, 2018, paste, sep="-"), sep="-")
names(gam_2018_tmax_st)
```


# Stack All Gambia Tmax values
```{r}
gam_2000_2018_tmax_st <- stack(gam_2000_tmax_st, gam_2001_tmax_st, gam_2002_tmax_st,
                               gam_2003_tmax_st, gam_2004_tmax_st, gam_2005_tmax_st,
                               gam_2006_tmax_st, gam_2007_tmax_st, gam_2008_tmax_st,
                               gam_2009_tmax_st, gam_2010_tmax_st, gam_2011_tmax_st,
                               gam_2012_tmax_st, gam_2013_tmax_st, gam_2014_tmax_st,
                               gam_2015_tmax_st, gam_2016_tmax_st, gam_2017_tmax_st,
                               gam_2018_tmax_st)

names(gam_2000_2018_tmax_st)
```

# Extract Values 2000-2018
```{r extracting max temp from 2000-2018}
# Whole of Gambia
extract_whole_gmb <- raster::extract(gam_2000_2018_tmax_st, extent(gam_regions)) %>%
  data.frame()

head(extract_whole_gmb)

# Upper River
extract_upper <- raster::extract(gam_2000_2018_tmax_st, extent(gam_regions_upper)) %>%
  data.frame()

head(extract_upper)

# Central River
extract_central <- raster::extract(gam_2000_2018_tmax_st, extent(gam_regions_central)) %>%
  data.frame()

head(extract_central)

# Lower River

extract_lower <- raster::extract(gam_2000_2018_tmax_st, extent(gam_regions_lower)) %>%
  data.frame()

head(extract_lower)
```


# Average Monthly Maximum Temperature
## The Whole of Gambia
```{r whole of gambia temp assessment}
# The whole of Gambia
extract_whole_gmb_tmax <- extract_whole_gmb %>%
  summarise_all(list(MIN=min, MEAN=mean, MAX=max), na.rm = T)

extract_whole_gmb_tmax 

# Minimums
extract_whole_gmb_tmax_min <- extract_whole_gmb_tmax %>%
  select(ends_with("MIN")) %>% t() %>% data.frame() %>% rownames_to_column() %>%
  rename(date = "rowname", temp =".")%>%
  mutate(date = str_extract(date, "[A-Za-z]{3}[.]\\d{4}")) %>%
  mutate(date = str_replace(date, "[.]", "-"))%>%
  mutate(date = mdy(date)) %>%
  mutate(month = month(date), year = year(date))

head(extract_whole_gmb_tmax_min)

# Maximums
extract_whole_gmb_tmax_max <- extract_whole_gmb_tmax %>%
  select(ends_with("MAX")) %>% t() %>% data.frame() %>% rownames_to_column() %>%
  rename(date = "rowname", temp =".")%>%
  mutate(date = str_extract(date, "[A-Za-z]{3}[.]\\d{4}")) %>%
  mutate(date = str_replace(date, "[.]", "-"))%>%
  mutate(date = mdy(date)) %>%
  mutate(month = month(date), year = year(date))

head(extract_whole_gmb_tmax_max)

# Mean
extract_whole_gmb_tmax_mean <- extract_whole_gmb_tmax %>%
  select(ends_with("MEAN")) %>% t() %>% data.frame() %>% rownames_to_column() %>%
  rename(date = "rowname", temp =".") %>%
  mutate(date = str_extract(date, "[A-Za-z]{3}[.]\\d{4}")) %>%
  mutate(date = str_replace(date, "[.]", "-"))%>%
  mutate(date = mdy(date)) %>%
  mutate(month = month(date), year = year(date))

head(extract_whole_gmb_tmax_mean)


# Merge into one dataset for the whole of Gambia
extracted_gambia_full <- cbind(extract_whole_gmb_tmax_min[,c(1:4)],
                               extract_whole_gmb_tmax_max[,2],
                               extract_whole_gmb_tmax_mean[,2])
head(extracted_gambia_full)

# Rename
extracted_gambia_full <- extracted_gambia_full %>%
                          rename(min_max_temp = "temp",
                                 max_max_temp = "extract_whole_gmb_tmax_max[, 2]",
                                 mean_max_temp = "extract_whole_gmb_tmax_mean[, 2]") %>%
  # order columns
  select(date, month, year, min_max_temp, max_max_temp, mean_max_temp)

head(extracted_gambia_full)

# Write the csv to file to use in regression model in gambia_rainfall_temp RMD
write_csv(extracted_gambia_full, "extracted_gambia_full.csv")
```
## Upper River Gambia
```{r upper river region temp assessment}
# The Upper River region
extract_upper_tmax <- extract_upper %>%
  summarise_all(list(MIN=min, MEAN=mean, MAX=max), na.rm = T)

head(extract_upper_tmax )

# Minimums
extract_upper_tmax_min <- extract_upper_tmax %>%
  select(ends_with("MIN")) %>% t() %>% data.frame() %>% rownames_to_column() %>%
  rename(date = "rowname", temp =".")%>%
  mutate(date = str_extract(date, "[A-Za-z]{3}[.]\\d{4}")) %>%
  mutate(date = str_replace(date, "[.]", "-"))%>%
  mutate(date = mdy(date)) %>%
  mutate(month = month(date), year = year(date))

head(extract_upper_tmax_min)

# Maximums
extract_upper_tmax_max <- extract_upper_tmax %>%
  select(ends_with("MAX")) %>% t() %>% data.frame() %>% rownames_to_column() %>%
  rename(date = "rowname", temp =".")%>%
  mutate(date = str_extract(date, "[A-Za-z]{3}[.]\\d{4}")) %>%
  mutate(date = str_replace(date, "[.]", "-"))%>%
  mutate(date = mdy(date)) %>%
  mutate(month = month(date), year = year(date))

head(extract_upper_tmax_max)

# Mean
extract_upper_tmax_mean <- extract_upper_tmax %>%
  select(ends_with("MEAN")) %>% t() %>% data.frame() %>% rownames_to_column() %>%
  rename(date = "rowname", temp =".") %>%
  mutate(date = str_extract(date, "[A-Za-z]{3}[.]\\d{4}")) %>%
  mutate(date = str_replace(date, "[.]", "-"))%>%
  mutate(date = mdy(date)) %>%
  mutate(month = month(date), year = year(date))

head(extract_upper_tmax_mean)


# Merge into one dataset for the Upper River Gambia
extracted_gambia_upper <- cbind(extract_upper_tmax_min[,c(1:4)],
                               extract_upper_tmax_max[,2],
                               extract_upper_tmax_mean[,2])
head(extracted_gambia_upper)

# Rename
extracted_gambia_upper <- extracted_gambia_upper %>%
                          rename(min_max_temp = "temp",
                                 max_max_temp = "extract_upper_tmax_max[, 2]",
                                 mean_max_temp = "extract_upper_tmax_mean[, 2]") %>%
  # order columns
  select(date, month, year, min_max_temp, max_max_temp, mean_max_temp) %>%
  # Add region column
  mutate(region = "Upper River")

head(extracted_gambia_upper)

```


## The Central River Region of Gambia
```{r central river region temp assessment}
# The Central River Region of Gambia
extract_central_tmax <- extract_central %>%
  summarise_all(list(MIN=min, MEAN=mean, MAX=max), na.rm = T)

head(extract_central_tmax)

# Minimums
extract_central_tmax_min <- extract_central_tmax %>%
  select(ends_with("MIN")) %>% t() %>% data.frame() %>% rownames_to_column() %>%
  rename(date = "rowname", temp =".")%>%
  mutate(date = str_extract(date, "[A-Za-z]{3}[.]\\d{4}")) %>%
  mutate(date = str_replace(date, "[.]", "-"))%>%
  mutate(date = mdy(date)) %>%
  mutate(month = month(date), year = year(date))

head(extract_central_tmax_min)

# Maximums
extract_central_tmax_max <- extract_central_tmax %>%
  select(ends_with("MAX")) %>% t() %>% data.frame() %>% rownames_to_column() %>%
  rename(date = "rowname", temp =".")%>%
  mutate(date = str_extract(date, "[A-Za-z]{3}[.]\\d{4}")) %>%
  mutate(date = str_replace(date, "[.]", "-"))%>%
  mutate(date = mdy(date)) %>%
  mutate(month = month(date), year = year(date))

head(extract_central_tmax_max)

# Mean
extract_central_tmax_mean <- extract_central_tmax %>%
  select(ends_with("MEAN")) %>% t() %>% data.frame() %>% rownames_to_column() %>%
  rename(date = "rowname", temp =".") %>%
  mutate(date = str_extract(date, "[A-Za-z]{3}[.]\\d{4}")) %>%
  mutate(date = str_replace(date, "[.]", "-"))%>%
  mutate(date = mdy(date)) %>%
  mutate(month = month(date), year = year(date))

head(extract_central_tmax_mean)



# Merge into one dataset for the Central River Gambia
extracted_gambia_central <- cbind(extract_central_tmax_min[,c(1:4)],
                               extract_central_tmax_max[,2],
                               extract_central_tmax_mean[,2])
head(extracted_gambia_central)

# Rename
extracted_gambia_central <- extracted_gambia_central %>%
                          rename(min_max_temp = "temp",
                                 max_max_temp = "extract_central_tmax_max[, 2]",
                                 mean_max_temp = "extract_central_tmax_mean[, 2]") %>%
  # order columns
  select(date, month, year, min_max_temp, max_max_temp, mean_max_temp)%>%
  # Add region column
  mutate(region = "Central River")

head(extracted_gambia_central)

```

## The Lower Region of Gambia
```{r lower river region temp assessment}
# The Lower Region of Gambia
extract_lower_tmax <- extract_lower %>%
  summarise_all(list(MIN=min, MEAN=mean, MAX=max), na.rm = T)

extract_lower_tmax 

# Minimums
extract_lower_tmax_min <- extract_lower_tmax %>%
  select(ends_with("MIN")) %>% t() %>% data.frame() %>% rownames_to_column() %>%
  rename(date = "rowname", temp =".")%>%
  mutate(date = str_extract(date, "[A-Za-z]{3}[.]\\d{4}")) %>%
  mutate(date = str_replace(date, "[.]", "-"))%>%
  mutate(date = mdy(date)) %>%
  mutate(month = month(date), year = year(date))

head(extract_lower_tmax_min)

# Maximums
extract_lower_tmax_max <- extract_lower_tmax %>%
  select(ends_with("MAX")) %>% t() %>% data.frame() %>% rownames_to_column() %>%
  rename(date = "rowname", temp =".")%>%
  mutate(date = str_extract(date, "[A-Za-z]{3}[.]\\d{4}")) %>%
  mutate(date = str_replace(date, "[.]", "-"))%>%
  mutate(date = mdy(date)) %>%
  mutate(month = month(date), year = year(date))

head(extract_lower_tmax_max)

# Mean
extract_lower_tmax_mean <- extract_lower_tmax %>%
  select(ends_with("MEAN")) %>% t() %>% data.frame() %>% rownames_to_column() %>%
  rename(date = "rowname", temp =".") %>%
  mutate(date = str_extract(date, "[A-Za-z]{3}[.]\\d{4}")) %>%
  mutate(date = str_replace(date, "[.]", "-"))%>%
  mutate(date = mdy(date)) %>%
  mutate(month = month(date), year = year(date))

head(extract_lower_tmax_mean)


# Merge into one dataset for the Lower River Gambia
extracted_gambia_lower <- cbind(extract_lower_tmax_min[,c(1:4)],
                               extract_lower_tmax_max[,2],
                               extract_lower_tmax_mean[,2])
head(extracted_gambia_lower)

# Rename
extracted_gambia_lower <- extracted_gambia_lower %>%
                          rename(min_max_temp = "temp",
                                 max_max_temp = "extract_lower_tmax_max[, 2]",
                                 mean_max_temp = "extract_lower_tmax_mean[, 2]") %>%
  # order columns
  select(date, month, year, min_max_temp, max_max_temp, mean_max_temp) %>%
  # Add region column
  mutate(region = "Lower River")

head(extracted_gambia_lower)

```


